---
title: "Class10_miniproj"
author: "Byanca (16688058)"
date: "10/28/2021"
output: html_document
---

#Get the data
Here we look at 538 Halloween candy. We take the CSV 

```{r}
url <-"https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy <- read.csv(url,row.names=1)
head(candy, n=5)
```
>Q1.

```{r}
nrow(candy)
```
>Q2.

```{r}
sum(candy$fruity)
```

#2. What is your favorite candy?

```{r}
candy["Snickers",]$winpercent
```

>Q4. 

```{r}
candy["Kit Kat",]$winpercent
```
>Q5.

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```



```{r}
library("skimr")
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
      Win percent
      
>Q7. What do you think a zero and one represent for the candy$chocolate column?

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

>Q9. Is the distribution of winpercent values symmetrical?
No

>Q10. Is the center of the distribution above or below 50%?
Below

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?
Chocolate candy is higher ranked than fruity candy

First we need to final all the chocolate candy rows in my 'candy' data

```{r}
inds<- as.logical(candy$chocolate)
candy[inds,]$winpercent 
```

```{r}
inds<- as.logical(candy$chocolate)
chocolate<- candy[inds,]$winpercent 

inds.fruit<- as.logical(candy$fruity)
fruity<- candy[inds.fruit,]$winpercent 

mean(chocolate)
mean(fruity)
```
>Q12. Is this difference statistically significant?
Yes

```{r}
t.test(chocolate, fruity)
```

#3. Overall Candy Rankings

>Q13. What are the five least liked candy types in this set?
Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, Jawbusters

```{r}
head(candy[order(candy$winpercent),], n=5)
```



>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent),], n=)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy)+
  aes(winpercent, rownames(candy))+
  geom_col()
```

Make it better
>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy)+
  aes(winpercent, reorder(rownames(candy), winpercent))+
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "red"
my_cols[as.logical(candy$bar)] = "orange"
my_cols[as.logical(candy$fruity)] = "brown"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
>Q17. What is the worst ranked chocolate candy?
Sixlets

>Q18. What is the best ranked fruity candy?
Starburst

#4. Taking a look at pricecenter

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
Reese's Minatures

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
a. Nik L Nip, Nestle Smarties, Ring pop, Hershey's Krackel. Hershey's Milk Chocolate
b. Nik L Nip  

#5. Exploring the correlation structure

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```
>Q22. What two variables are anti-correlated?
chocolate and fruity

>Q23. What two varibles are most positvely correlated?
chocolate and winpercent

#6. Principal Component Analysis
```{r}
pca<- prcomp(candy, scale=TRUE)
summary(pca)
```
 
```{r}
plot(pca$x[,1:2], col=my_cols)
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
# Make a new data-frame with our PCA results and candy data

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```



```{r}
library(plotly)
```
```{r}
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
Fruity, hard, pluribus